Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-10-10 paper/galsize_paper.tex
Modified abstract

\title[Forward Modeling Galaxy Size]
{Forward Modeling Galaxy Size in a Cosmological Context}

@@

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius of galaxies, $\rhalf.$ Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. Smaller galaxies cluster stronger relative to larger galaxies of the same stellar mass, a new result. We use {\tt Halotools} to test a collection of forward models of galaxy size to identify the qualitative ingredients needed to reproduce the observed clustering. We show that the $\rhalf-$dependence of galaxy clustering is largely driven by centrals being larger than satellite galaxies of the same halo mass. Models in which $\rhalf$ is determined by stellar mass $\mstar$ exhibit qualitatively discrepant clustering properties from SDSS galaxies. Models where $\rhalf$ is linearly proportional to halo virial radius $\rvir$ at the time of peak halo mass are much more successful, provided that scatter in $\rhalf$ is significantly correlated with the halo scale radius $R_{\rm s}$ at that time. Together with the result that stellar mass stripping of satellites has only a mild impact on $\rhalf-$dependent clustering, this suggests that the relative size of centrals and satellites is already in place at the time of satellite infall, and supports the notion that $L_{\ast}$ galaxy and halo profiles co-evolve across many Gyr of cosmic time.
\end{abstract}

>>>>

\title[The Relative Sizes of Centrals and Satellites]
{Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies}

@@

\begin{abstract}
We place empirical constraints on the connection between dark matter halos and galaxy half-light radii, $\rhalf.$ Using the SDSS, we show that smaller galaxies cluster much more strongly than larger galaxies at fixed stellar mass.  Using Halotools to forward-model the observations, we find that the clustering signal generically requires satellite galaxies to be smaller than central galaxies of the same halo mass.  We present a simple empirical model consistent with the clustering results, in which galaxy size is proportional to halo virial radius at the time of peak halo mass, and the scatter is strongly correlated with halo scale radius.  Models in which galaxy size is related to stellar mass only or that have uncorrelated scatter in galaxy size both fail the clustering test, even considering effects of satellite stripping and orphan galaxies.  These results suggest that the relative size of centrals and satellites is predetermined at the time of satellite infall, and supports the notion that $L_\ast$ galaxy and halo profiles co-evolve across many Gyr of cosmic time.
\end{abstract}
